### 2022-07-12 src/docs_paper/appendix.Rmd

---
title: Appendix to "Models for spatial structure in small-area estimation"
subtitle: |
  Corresponding author: Adam Howes (`ath19@ic.ac.uk`)
output:
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

<!-- So that the tables and figures in this appendix are prefixed by "A." -->
\setcounter{table}{0}  
\renewcommand{\thetable}{A.\arabic{table}} 
\setcounter{figure}{0} 
\renewcommand{\thefigure}{A.\arabic{figure}}

\newpage

# Implementation details for the Besag model \label{app:freni}

Here we briefly review three best practices for using the Besag model, scaling, singletons, and constraints, as recommended by @freni2018note:

1. \textit{Scaling}
The structure matrix $\mathbf{R}$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
\begin{equation}
    \sigma^2_{\text{GV}}(\mathbf{R}) = \prod_{i = 1}^n (R^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (R^-_{ii}) \right),
\end{equation}
equal to one, by replacing $\mathbf{R}$ with $\mathbf{R}^\star = \mathbf{R} / \sigma^2_{\text{GV}}(\mathbf{R})$.
As the diagonal elements $R^-_{ii}$ correspond to marginal variances, the generalised variance gives a measure of the average marginal variance.
However, this measure, introduced by @sorbye2014scaling, ignores off-diagonal entries and more broadly any measure of typical variance could be used.
Scaling mitigates the influence of the adjacency graph on the variance of $\bphi$.
Allowing the variance to be controlled by $\tau_\phi$ alone is important as it allows for consistent, interpretable prior selection.
When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $\mathbf{R}$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
As such, each connected component should be scaled independently to have generalised variance one giving $\mathbf{R}^\star_I = \mathbf{R}_I / \sigma^2_{\text{GV}}(\mathbf{R}_I)$ where $\mathbf{R}_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.

2.
\textit{Singletons}
When one of the connected components is a single area, known either as a singleton or an island, the probability density $\exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right)$ has no dependence on $\phi_i$.
This is equivalent to using an improper prior $p(\phi_i) \propto 1$ and can be avoided by setting each singleton to have independent Gaussian noise $p(\phi_i) \sim \mathcal{N}(0, 1)$.

3.
\textit{Constraints}
To avoid confounding of the spatial random effects with the intercept, it is recommended to place a sum-to-zero constraint on each non-singleton connected component.
In other words, for each $|I| > 1$ that $\sum_{i \in I} \phi_i = 0$.

# References {#references .unnumbered}