### 2022-12-13 src/1_checks/inla-mcmc.Rmd
Consolidate 1_checks

---
title: "Check R-INLA against MCMC"
author:
  - name: "Adam Howes"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task**
---

```{r}
#' Checking that R-INLA and Stan produce similar results
#' Would using Stan (slower) for the models below rather than R-INLA (faster) change anything?

#' Example data, should repeat this for others
data <- readRDS("depends/data_iid_grid.rds")
sf <- data[[1]]$sf

#' Constant
fit_inla <- constant_inla(sf)
fit_stan <- constant_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' IID
fit_inla <- iid_inla(sf)
fit_stan <- iid_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' Besag
fit_inla <- besag_inla(sf)
fit_stan <- besag_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' BYM2
fit_inla <- bym2_inla(sf)
fit_stan <- bym2_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' TODO: These plots could be improved, ggplot2, a line through x = y
#' Could also check how things look for more than one dataset
```