### 2022-12-13 src/1_checks/plausible-lengthscale.Rmd

---
title: "Average distances between centroids for plausible lengthscale choice"
author:
  - name: "Adam Howes"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task**
---

```{r}
grid <- readRDS("depends/geometry_grid.rds")
civ <- readRDS("depends/geometry_civ.rds")
tex <- readRDS("depends/geometry_tex.rds")

grid_distances <- c(centroid_distance(grid))
min(grid_distances[grid_distances > 0])

civ_distances <- c(centroid_distance(civ))
min(civ_distances[ci_distances > 0])

tex_distances <- c(centroid_distance(tex))
min(tex_distances[tex_distances > 0])

cowplot::plot_grid(
  graph_length_plot(grid),
  ggplot() + geom_histogram(data = data.frame(x = grid_distances), aes(x = x))
)

cowplot::plot_grid(
  graph_length_plot(civ),
  ggplot() + geom_histogram(data = data.frame(x = civ_distances), aes(x = x))
)

cowplot::plot_grid(
  graph_length_plot(tex),
  ggplot() + geom_histogram(data = data.frame(x = tex_distances), aes(x = x))
)
```