### 2023-05-31 src/docs_paper/appendix.Rmd
Will Adam ever finish writing this paper? Stay tuned to find out (I hope so)

---
title: Appendix to "Understanding models for spatial structure in small-area estimation"
author:
  - Adam Howes^[Department of Mathematics, Imperial College London]
  - Jeffrey W. Eaton^[Department of Epidemiology, Harvard University]
  - Seth R. Flaxman^[Department of Computer Science, Oxford University]
output:
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

```{r echo = FALSE}
options(scipen = 100)
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 320,
  cache = TRUE,
  out.width = "95%",
  fig.align = 'center'
)

comma <- function(x) formatC(x, format = "d", big.mark = ",")
```

<!-- So that the tables and figures in this appendix are prefixed by "A." -->
\setcounter{table}{0}  
\renewcommand{\thetable}{A.\arabic{table}} 
\setcounter{figure}{0} 
\renewcommand{\thefigure}{A.\arabic{figure}}

\clearpage

# Comparison of AGHQ to other inference methods \label{app:aghq-nuts}

\newpage

# Implementation details for the Besag model \label{app:freni}

Here we briefly review three best practices for using the Besag model, scaling, singletons, and constraints, as recommended by @freni2018note:

## Scaling

The structure matrix $\mathbf{R}$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
\begin{equation}
    \sigma^2_{\text{GV}}(\mathbf{R}) = \prod_{i = 1}^n (\mathbf{R}^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (\mathbf{R}^-_{ii}) \right),
\end{equation}
equal to one, by replacing $\mathbf{R}$ with $\mathbf{R}^\star = \mathbf{R} / \sigma^2_{\text{GV}}(\mathbf{R})$.
As the diagonal elements $R^-_{ii}$ correspond to marginal variances, the generalised variance gives a measure of the average marginal variance.
However, this measure, introduced by @sorbye2014scaling, ignores off-diagonal entries and more broadly any measure of typical variance could be used.
Scaling mitigates the influence of the adjacency graph on the variance of $\bphi$.
Allowing the variance to be controlled by $\tau_\phi$ alone is important as it allows for consistent, interpretable prior selection.
When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $\mathbf{R}$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
As such, each connected component should be scaled independently to have generalised variance one giving $\mathbf{R}^\star_I = R_I / \sigma^2_{\text{GV}}(R_I)$ where $R_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.

## Singletons

When one of the connected components is a single area, known either as a singleton or an island, the probability density $\exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right)$ has no dependence on $\phi_i$.
This is equivalent to using an improper prior $p(\phi_i) \propto 1$ and can be avoided by setting each singleton to have independent Gaussian noise $p(\phi_i) \sim \mathcal{N}(0, 1)$.

## Constraints

To avoid confounding of the spatial random effects with the intercept, it is recommended to place a sum-to-zero constraint on each non-singleton connected component.
In other words, for each $|I| > 1$ that $\sum_{i \in I} \phi_i = 0$.

\clearpage

# Further results for the simulation study

```{r results="asis"}
#' Some of the lines should not be included
lines_remove <- c(1, 3, 4, 5)
cat(readLines("depends/crps-table-rho.txt")[-lines_remove], sep = "\n")
```

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=1]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-1}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=2]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-2}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=3]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-3}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=4]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-4}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=5]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-5}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=6]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-6}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=7]{depends/crps-boxplot-rho.pdf}}
\label{fig:crps-boxplot-rho-7}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=1]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-1}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=2]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-2}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=3]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-3}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=4]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-4}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=5]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-5}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=6]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-6}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=7]{depends/histogram-ecdf.pdf}}
\label{fig:histogram-ecdf-7}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=1]{depends/crps-maps.pdf}}
\label{fig:crps-maps-1}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=2]{depends/crps-maps.pdf}}
\label{fig:crps-maps-2}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=3]{depends/crps-maps.pdf}}
\label{fig:crps-maps-3}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=4]{depends/crps-maps.pdf}}
\label{fig:crps-maps-4}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=5]{depends/crps-maps.pdf}}
\label{fig:crps-maps-5}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=6]{depends/crps-maps.pdf}}
\label{fig:crps-maps-6}
\end{figure}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{\includegraphics[page=7]{depends/crps-maps.pdf}}
\label{fig:crps-maps-7}
\end{figure}

\clearpage

# Further results for the HIV study

\clearpage

# References {#references .unnumbered}